Bootstrap Confidence Intervals for Sharp Regression Discontinuity Designs with the Uniform Kernel

### 2016-02-29 main_paper.tex
Make small edits to the paper's introduction

\newpage
\section{Introduction}
Regression Discontinuity (RD) designs have emerged in the last decade as an
important and popular research design strategy for analyzing the causal impact
of policies and interventions in several fields of the social sciences,
including economics, political science, public policy, and sociology, among
others. This strategy exploits the fact that many programs use somewhat
arbitrary thresholds to determine whether or not to provide a ``treatment,''
usually relying on some type of score, which this literature calls a ``running
variable.'' In its basic version, individuals or groups with score above a
certain threshold receive treatment while the ones below that cutoff are left
untreated.

The identification of the treatment effect at the threshold is then based on
comparing treated and untreated units at the cutoff. Since the observed
characteristics of these subjects are very similar, similarity of their
unobserved characteristics is often credible as well, and differences in the
outcomes between the treated and untreated can be plausibly attributed to the
treatment alone. As a practical matter this involves comparing units within a
bandwidth just above and just below the threshold.

The RD design strategy was introduced by \cite{thistlethwaite1960} to study
educational outcomes, and many of its first recent applications in economics
were to estimate the effects of other educational policies: evaluating the
impact of investments in school facilities, class sizes, remedial education,
early childhood education, and financial aid effects on student achievement and
later outcomes, for example.\footnote{%
  See, for example, \cite{vdk2002}, \cite{jacoblefgren2004}, \cite{ludwig2007},
  \cite{urquiola2009}, \cite{cellini2010}} %
But the underlying identification strategy has proven to apply much more widely,
and RD has been used in health economics\footnote{%
  \cite{Card2009,barreca2011saving}}; %
political science, providing evidence regarding incumbency advantage in
elections, electoral enfranchising and policy changes associated with the
introduction of electronic voting, strategic voting behavior, and local media
effects and ad expenditure in presidential elections in the U.S.\footnote{%
  \cite{lee2008}, \cite{Caughey2011}, \cite{keele2014geographic},
  \cite{erikson2015}, \cite{Fujiwara2011,Fujiwara2015}}, %
among other fields. \cite{imbens2008} and \cite{lee2010} provide recent
overviews of this literature with many more examples.

In these studies, identification typically occurs exactly at the cutoff, so the
treatment effect is typically estimated by fitting separate local linear models
above and below the cutoff, then extrapolating the model to the exact point of
discontinuity. The difference between the estimated effects at that point is
taken to be an estimate of the treatment effect. As a practical matter, the key
econometric issue is determining the bandwidth for the local linear models.  One
very popular choice is the bandwidth estimator proposed by \cite{IK} and
extended by \cite{calonico2014}, which minimizes the Asymptotic Mean Squared
Error of difference in the models' point estimators at the cutoff.

But, as \cite{calonico2014}\footnote{%
  We will refer to \cite{calonico2014} frequently, so we will reference it as
  ``CCT'' for the rest of the paper.} %
point out, the \cite{IK} bandwidth estimator has the serious drawback that,
although it is AMSE-optimal for point estimation, it produces invalid
\emph{confidence intervals} and hypothesis tests. The AMSE-optimal bandwidth
converges to zero slowly, so the remaining bias term is large enough to affect
the asymptotic distribution of the estimator. In that case the usual ``naive''
confidence intervals for the RD treatment effects may be biased and have
coverage well below their nominal level. CCT then make two contributions: first,
they propose an estimator of the bias resulting from undersmoothing and provide
a treatment effect estimator that remains asymptotically unbiased even when the
bandwidth converges to zero slowly, at the AMSE-optimal rate.  Second, since
this bias-correction term contributes to the asymptotic variance of the
resulting treatment effect estimator, CCT also provide a new formula for its
asymptotic variance. The resulting confidence intervals have accurate coverage
even when the naive RD interval does not.

@@

Our approach exploits CCT's theoretical insight through a new residual
bootstrap. In particular, we propose, after estimating the local linear model as
usual, estimating a local second order polynomial and generating bootstrap
datasets by resampling the residuals of that polynomial. Since the second order
polynomial is the true Data Generating Process (DGP) in the bootstrapped data, its estimate of the
treatment effect is the true value of the treatment effect under the
distribution induced by this bootstrap. The bias of the linear model is known
under this distribution and can be calculated by averaging the error of the
linear model's estimates across many bootstrap replications. This approach is
described in detail by our Algorithm~\ref{Alg1} and the resulting bias corrected
estimator is shown to be asymptotically normal with mean zero in our
Theorem~\ref{T1} under AMSE-optimal bandwidth rates.

Just as in CCT, this bias correction introduces additional variability. However,
the second order polynomial again adequately estimates the features of the true
DGP that are necessary for estimating that additional variability. So we propose
an iterated bootstrap procedure \citep{hall1988}: use the second order
polynomial residual bootstrap to produce many bootstrap replications of the bias
corrected estimator, and then use the resulting bootstrap distribution to
produce confidence intervals. This procedure, which requires bootstrapping the
datasets produced by an initial bootstrap procedure, is described in Algorithm~\ref{Alg2}, and the resulting confidence intervals are shown to be asymptotically valid in Theorem~\ref{T2}.

This bootstrap procedure can offer some advantages over analytical methods in
this setting. In particular, both this paper and CCT assume that the
observations are generated independently of each other; extending these bias
correction methods to other forms of dependence is relatively straightforward
for the bootstrap but can be substantially more complicated for analytical
corrections, which have to be explicitly derived by the researcher. Similarly,
although we only provide results for the local linear model in this paper, it is
trivial to implement this procedure for higher order polynomials, for
covariate-adjusted estimators (\citealp{frolich2007}, \citealp{calonico2015}),
or for other local smoothers. \citep{loader2006} However, in this paper we focus
on the baseline case of sharp RD with a local linear model and uniform
kernel. Extensions to fuzzy RD designs\footnote{%
  ``Fuzzy'' regression discontinuity design \citep{HTV2001}, as opposed to
  ``sharp'' RD, describes situations where the probability of treatment changes
  discontinuously at a known threshold, but by less than 100\%. Then there are
  treated and untreated subjects above and below the discontinuity but the
  treatment effect remains identified.} %
and nonuniform kernels, which require nontrivial changes to the underlying
bootstrap algorithm, as well as developments to address cross sectional
dependence, are the subject of ongoing research.

>>>>

\newpage
\section{Introduction}
Regression Discontinuity (RD) designs have emerged in the last decade as an
important and popular research design strategy for analyzing the causal impact
of policies and interventions in several fields of the social sciences,
including economics, political science, public policy, and sociology.
This research strategy exploits the fact that many programs use somewhat
arbitrary thresholds to determine whether or not to provide a ``treatment,''
usually relying on some type of score, which this literature calls a ``running
variable.'' In its basic version, individuals or groups with score above a
certain threshold receive treatment while the ones below that cutoff are left
untreated.
The identification of the treatment effect at the threshold is then based on
comparing treated and untreated units at the cutoff. Since the observed
characteristics of these subjects are very similar, similarity of their
unobserved characteristics is often credible as well, and differences in the
outcomes between the treated and untreated can be plausibly attributed to the
treatment alone. As a practical matter this involves comparing units within a
bandwidth just above and just below the threshold.

The RD design strategy was introduced by \cite{thistlethwaite1960} to study
educational outcomes, and many of its first recent applications in economics
were to estimate the effects of other educational policies: evaluating the
impact of investments in school facilities, class sizes, remedial education,
early childhood education, and financial aid effects on student achievement and
later outcomes, for example.\footnote{%
  See, for example, \cite{vdk2002}, \cite{jacoblefgren2004}, \cite{ludwig2007},
  \cite{urquiola2009}, \cite{cellini2010}} %
But the underlying identification strategy has proven to apply much more widely
and RD has been used in health economics;\footnote{%
  \cite{Card2009,barreca2011saving}} %
political science, providing evidence regarding incumbency advantage in
elections, electoral enfranchising and policy changes associated with the
introduction of electronic voting, strategic voting behavior, and local media
effects and ad expenditure in presidential elections in the U.S.\footnote{%
  \cite{lee2008}, \cite{Caughey2011}, \cite{keele2014geographic},
  \cite{erikson2015}, \cite{Fujiwara2011,Fujiwara2015}}, %
among other fields. \cite{imbens2008} and \cite{lee2010} provide recent
overviews of this literature with many more examples.

In these studies, identification occurs exactly at the cutoff, so the
treatment effect is typically estimated by fitting separate local linear models
above and below the cutoff, then extrapolating the models to the exact point of
discontinuity. The difference between the estimated effects at that point is
taken to be an estimate of the treatment effect. As a practical matter, the key
econometric issue is determining the bandwidth for the local linear models.  One
very popular choice is the bandwidth estimator proposed by \cite{IK} and
extended by \cite{calonico2014}, which minimizes the Asymptotic Mean Squared
Error of difference in the models' point estimators at the cutoff.

As \cite{calonico2014}\footnote{%
  We will refer to \cite{calonico2014} frequently, so we will reference it as
  ``CCT'' for the rest of the paper.} %
point out, however, the \cite{IK} bandwidth estimator has the serious drawback
that, although it is AMSE-optimal for point estimation, it produces invalid
confidence intervals and hypothesis tests. The AMSE-optimal bandwidth
converges to zero slowly, so the remaining bias term is large enough to affect
the asymptotic distribution of the estimator. In that case the usual ``naive''
confidence intervals for the RD treatment effects may be biased and have
coverage well below their nominal level.
CCT propose an estimator of the bias resulting from undersmoothing and provide a
treatment effect estimator that remains asymptotically unbiased even when the
bandwidth converges to zero slowly, at the AMSE-optimal rate. They also show
that this bias-correction term contributes to the asymptotic variance of the
resulting treatment effect estimator and provide a new formula for the
asymptotic variance of the bias-corrected estimator. The resulting confidence
intervals have accurate coverage even when the naive RD interval does not.

@@

Our approach exploits CCT's theoretical insight through a new residual
bootstrap. In particular, we propose estimating the local linear model as usual,
then estimating a local second order polynomial and generating bootstrap
datasets by resampling the residuals of that polynomial. Since the second order
polynomial is the true Data Generating Process (DGP) for the bootstrapped data,
its estimate of the treatment effect is the true value of the treatment effect
under the distribution induced by this bootstrap. The bias of the linear model
is therefore known
under this distribution and can be calculated by averaging the error of the
linear model's estimates across many bootstrap replications. This approach is
described in detail by our Algorithm~\ref{Alg1} and the resulting bias corrected
estimator is shown to be asymptotically normal with mean zero in our
Theorem~\ref{T1} under AMSE-optimal bandwidth rates.

Just as in CCT, our bias correction step introduces additional
variability. However, the second order polynomial again adequately estimates the
features of the true DGP that are necessary for estimating and accomodating that
additional variability. So we propose
an iterated bootstrap procedure \citep{hall1988}: use the second order
polynomial residual bootstrap to produce many bootstrap replications of the bias
corrected estimator, and then use the resulting bootstrap distribution to
produce confidence intervals. This procedure, which requires bootstrapping the
datasets produced by an initial bootstrap procedure, is described in Algorithm~\ref{Alg2}, and the resulting confidence intervals are shown to be asymptotically valid in Theorem~\ref{T2}.

This bootstrap procedure offers some advantages over analytical methods.  In
particular, both this paper and CCT assume that the observations are generated
independently of each other; however, extending these bias correction methods to
other forms of dependence is relatively straightforward for the bootstrap but
can be substantially more complicated for analytical
corrections, which have to be explicitly derived by the researcher. Similarly,
although we only provide results for the local linear model in this paper, it is
trivial to implement this procedure for higher order polynomials, for
covariate-adjusted estimators (\citealp{frolich2007}, \citealp{calonico2015}),
or for other local smoothers. \citep{loader2006} However, in this paper we focus
on the baseline case of sharp RD with a local linear model and uniform
kernel. Extensions to fuzzy RD designs\footnote{%
  ``Fuzzy'' regression discontinuity design \citep{HTV2001}, as opposed to
  ``sharp'' RD, describes situations where the probability of treatment changes
  discontinuously at a known threshold, but by less than 100\%. Then there are
  treated and untreated subjects above and below the discontinuity but the
  treatment effect remains identified.} %
and nonuniform kernels, which require nontrivial changes to the underlying
bootstrap algorithm, as well as developments to address cross sectional
dependence, are the subject of ongoing research.
